# R Data Structures


## Using Simple Data Types
## Integer 

```{r}
x <- 32
```

```{r}
x <- 32
32 -> x
```

## Numeric
```{r}
x <- 4.67
y <- 8.58
```

## Char

```{r}
myChar <- "a"
myName <- "John"
class(myChar)
class(myName)
```



## Logical

```{r}
x <- 1
y <- 2
x < y
```

## Complex
```{r}
x <- 2
y <- 3
z <- complex(real = x, imaginary = y)
z
```

## Creating datasets with vectors

## Creating simple vectors
```{r}
myFirstVector <- c (2, 14, 6, 17, 18, 9, 15, 25, 7, 11)
mean(myFirstVector)
n <- rep(2, 10) 
n <- rep(c(2, 5), 10)
seq(20)
seq(0, 20, 2)
seq(from = 0, to = 20, by = 2)
?seq
seq(from = 1, to = 1, by = ((to - from)/(length.out - 1)), length.out = NULL, along.with = NULL, ...) 
myCharacterVector <- c("a", "b", "c", "d", "e")
myStringVector <- c("apple", "grape", "pomegranate", "orange", "banana")
myCharacterVector <- c("a", "b", "c", "d", "e", 1, 2)
myCharacterVector

myStringVector[1]
myStringVector[2]


```

## Performing basic mathematic operations on vectors 

## Manipulating vectors with functions

## Storing characters and strings with vectors

## Matrices

### Creating a simple matrix
```{r}
myFirstMatrix <- matrix(1:20, 4, 5)
myFirstMatrix
myFirstMatrix <- matrix(1:20, nrow = 4, ncol = 5)
myFirstMatrix
myFirstMatrix <- matrix(1:20, nrow = 5, ncol = 4)
myFirstMatrix
myFirstMatrix <- matrix(1:20, 4, 5)
myFirstMatrix <- matrix(1:20, nrow = 4, ncol = 5)
myFirstMatrix <- matrix(1:20, 5, 4)
myFirstMatrix <- matrix(1:20, nrow = 5, ncol = 5)
```

### Combining vectors to create a matrix
```{r}
vector1 <- c(1:10)
vector2 <- c(101:110)
vector3 <- c(6:15)

matrix1 <- rbind(vector1, vector2, vector3)
matrix1

matrix2 <- cbind(vector1, vector2, vector3)
matrix2

colnames(matrix2) <- c("column1", "column2", "column3")
rownames(matrix2) <- c(1:10)
matrix2

matrix5 <- matrix(rep(1:10), 2, 5)
matrix5
matrix6 <- matrix(rep(11:25), 3, 5)
matrix6
matrix7 <- rbind(matrix5, matrix6)
matrix7
matrix8 <- matrix6[, -1]
matrix9 <- cbind(matrix5, matrix8)

```


### Combining matrices
```{r}
install.packages("gdata", dependencies=TRUE)
library(gdata)
matrix2
dim(matrix2)
matrix2
matrix3 <- matrix2[-10, ]
matrix3
dim(matrix3)
matrix2bind3<- cbindX(matrix2, matrix3)
```

### Selecting and extracting elements 

```{r}
matrix2[1, 1]
matrix2[7, 3]
matrix2[1, ]
matrix2[, 3]
```

## Data frames
```{r}
dataFrame3 <- as.data.frame(matrix3)
View(dataFrame3)
setwd("C:/Dropbox/5277OS - Learning Scientific Computing with R/5277os")
worms <- read.table("./worms.txt", header=T)
names(worms)
worms[2, 2]
worms[2, ]
wormsSoilPHOrdered <- worms[order(worms$Soil.pH), ]
detach (worms)
wormsDampTrue <- worms[worms$Damp==T, ]
wormsRows2To10Removed <- worms[-(2:10),]
wormsNAremoved <- na.exclude(worms)
rbind.fill(mtcars[c("mpg", "wt")], mtcars[c("wt", "cyl")])
```


## Apply functions family
### Apply
```{r}
matrix8 <- matrix(c(1:20, 41:60), nrow = 10, ncol = 2)
View(matrix8)
apply(matrix8, 1, mean)
apply(matrix8, 2, mean)
apply(matrix8, 1:2, function(x) {sqrt(x) } )
```

### Lapply
```{r}
myList <- list(a = 21:40, b = 31:50)
lapply(myList, sqrt)
```


### Sapply
```{r}
sapply(myList, sqrt)
```
## Plyr package
```{r}
set.seed(12)
d <- data.frame(year = rep(2000:2002, each = 3), count = round(runif(9, 0, 20)))
d <- data.frame(year = rep(2000:2002, each = 3), count = round(runif(9, 0, 20)))
print(d)
library(plyr)
ddply(d, "year", function(x) {
      mean.count <- mean(x$count)
      sd.count <- sd(x$count)
      cv <- sd.count/mean.count
      data.frame(cv.count = cv)
     }
  )



ddply(d, "year", summarise, mean.count = mean(count))

ddply(d, "year", transform, total.count = sum(count))

ddply(d, "year", mutate, mu = mean(count), sigma = sd(count), cv = sigma/mu)
```
